---
title: "Getting your blank maps"
output: 
    html_document:
      df_print: tibble
      toc: yes
      toc_float: yes
---

We need to have "bank maps" to color in. Two sources:

1. Dowloaded file.
2. Provided through an R-package [focus].

Points of attention:

* We want sf-objects. Frequently already right format, or convert using [`st_as_sf()`](https://r-spatial.github.io/sf/reference/st_as_sf.html).
* Reduce large map to what you need: filter, crop, or limit.
* Pay attention to included identifier.

```{r, message=FALSE, warning=FALSE}
# load general packages used below
library(ggplot2)
library(sf)
library(tmap)
library(mapview)
```


# Get maps from downloaded file

Formats: Shapefiles, GeoJSON, KML, etc.

Recommended: 

* Use `readOGR()` in the rgdal-package: hard to find a format not covered.
* Convert to simple features-object using 

Semi-random example: [download the open-data spatial boundaries](https://geo.onroerenderfgoed.be/downloads) of the management plans of the [Flemish organization for Immovable Heritage](https://en.wikipedia.org/wiki/Flemish_organization_for_Immovable_Heritage).

```{r, message=FALSE, warning=FALSE}
library(rgdal)
heritage <- readOGR('data/heritage_plans', layer = 'heritage_plans')
heritage <- st_as_sf(heritage)
```


```{r}
qtm(heritage)
```

```{r}
mapview(heritage)
```
<br>
<br>


# Get maps through R-packages


## Belgium

[BelgiumMaps.StatBel](http://www.bnosac.be/index.php/blog/55-belgiummaps-statbel-r-package-with-administrative-boundaries-of-belgium) by Jan Wijfels ([bnosac](http://www.bnosac.be/)): convenient R package bundeling spatial open data on Belgian administrative boundaries.

Load package (`library(BelgiumMaps.StatBel)`) and then use `data()` to load the spatial data for the administrative level you need: 

* BE_ADMIN_SECTORS: statistical sector / statistische sector
* BE_ADMIN_MUNTY: municipality / gemeente
* BE_ADMIN_DISTRICT: district / arrondissement
* BE_ADMIN_PROVINCE: province / provincie
* BE_ADMIN_REGION: region / regio
* BE_ADMIN_BELGIUM: country / land

Important: the data always contains a variable/column with the [NIS-code](https://en.wikipedia.org/wiki/NIS_code), named "CD_[level]_REFNIS". E.g. CD_PROV_REFNIS, CD_RGN_REFNIS, CD_MUNTY_REFNIS. With NIS-codes in your data, you can merge on each level. It also contains NUTS-codes (not demonstrated).

```{r}
library(BelgiumMaps.StatBel)
data("BE_ADMIN_PROVINCE") # load spatial object for provincial level
provinces <- st_as_sf(BE_ADMIN_PROVINCE) # convert to sf-object
```


```{r}
qtm(provinces) # plot with tmap
```



```{r}
data("BE_ADMIN_MUNTY")
munip <- st_as_sf(BE_ADMIN_MUNTY)
qtm(munip)
```


## Europe

The [eurostat](http://ropengov.github.io/eurostat/) package allows you to directly download, analyse and visualise data from [Eurostat](https://ec.europa.eu/eurostat) in R, including their blank maps of EU-member countries:

* Use the function `get_eurostat_geospatial()` to download.
* Choose the NUTS level: "0" (countries), "1" (regions, i.e. Flanders), "2" (sub-region, BE: provinces), "3" (sub-sub-region, BE: arrondissment).
* Choose the level of detail: "60" (1:60million), "20" (1:20million), "10" (1:10million), "01" (1:1million). More detail for more zoomed-in maps (longer download, but cached).
* Identifier: NUTS ("NUTS_ID").
* sf-object by default, no `st_as_sf()` needed.
* Warning: not EU-member (+ NO, TR, etc.), not on the map!

```{r, message=FALSE, warning=FALSE}
library(eurostat)
eu_nuts0 <- get_eurostat_geospatial(
  resolution = "60", # detail 
  nuts_level = "0") # NUTS 0-3

eu_nuts2 <- get_eurostat_geospatial(
  resolution = "60", # detail 
  nuts_level = "2") # NUTS 0-3
```


```{r}
qtm(eu_nuts0)
```

```{r}
qtm(eu_nuts2)
```




## World

Various R packages containing spatial data for the entire world, e.g.:

* tmap: load using `data("World").
* rworldmap (regular resolution) and complementary package rworldxtra (high resolution) 

```{r}
library(tmap)
data("World")
world_tmap <- st_as_sf(World)
qtm(world_tmap)
```

```{r, message=FALSE, warning=FALSE}
library(rworldmap)
#library(rworldxtra)

# load worldmap with resolution "coarse", "low", "less islands", "li", "high".
# for option "high" the additional package rworldxtra needs to be install, works the same.
world_worldmap <- getMap(resolution = "low")
world_worldmap <- st_as_sf(world_worldmap)
```

```{r, message=FALSE, warning=FALSE}
qtm(world_worldmap)
```


# Get rid of too much map

Common situation, three options:

1. **Filter** or select (un)wanted spatial data. 
2. **Crop** the map to the required area.
3. **Limit** the visible map area.

## Filter or select

Use a variable (originally in spatial dataset or added by you) to filter/select what you want from a larger map.

```{r}
data("BE_ADMIN_PROVINCE")
prov <- st_as_sf(BE_ADMIN_PROVINCE)
```

```{r}
qtm(prov)
```

```{r}
# filter out only the provinces in the region of Flanders, i.e. 
# where the region description variable "TX_RGN_DESCR_NL" is equal to "Vlaams Gewest"
prov.fl <- prov %>%
  filter(TX_RGN_DESCR_NL == 'Vlaams Gewest')
```

```{r}
qtm(prov.fl)
```


```{r, message=FALSE, warning=FALSE}
# filter out the South American countries from the world map
south_am <- world_worldmap %>% 
  filter(GEO3 == 'South America')

qtm(south_am)
```


## Limit visible area

Happens "at the end" when displaying the map, so depedent on what you use to display/plot. Example with tmap and ggplot:

```{r}
qtm(eu_nuts2, bbox = 'France')
```

```{r, message=FALSE, warning=FALSE}
ggplot(eu_nuts0) +
  geom_sf() +
  coord_sf(
    # limit map to 'mainland' EU
    xlim = c(2500000, 6000000), ylim =c(1500000, 5300000),
    crs = 3035)
```


## Crop spatial features

Use `st_crop()` from the sf package to crop a map to certain limits (remove rest)

```{r, message=FALSE, warning=FALSE}
eu <- st_crop(eu_nuts0, c(xmin=-10, xmax=45, ymin=36, ymax=71))
qtm(eu)
```




## Combine filter, crop, limit

```{r, message=FALSE, warning=FALSE}
# Filter out the Benelux based on country-names 
benelux <- world_worldmap %>%
  filter(NAME %in% c('Belgium', 'Netherlands', 'Luxembourg'))

# Plot Benelux and (exactly) limit map (decolonisation)
ggplot(benelux) +
  geom_sf() +
  coord_sf(xlim = c(3700000, 4300000),
           ylim =c(2800000, 3500000), crs = 3035)
```


