---
title: "Adding data"
output: 
    html_document:
      df_print: tibble
      toc: yes
      toc_float: yes
---

We have a (spatial) dataframe with the blank map-information, now we want to add our data-of-interest. Two steps:

1. Read in data-of-interst into R.
2. Join data-of-interest with the spatial dataframe.


```{r}
library(BelgiumMaps.StatBel)
library(tmap)
library(tmaptools)
library(dplyr)
library(readr)
library(readxl)
library(haven)
```

# Read in data-of-interest

Crucial: have a variable/column in your that contains the appropriate spatial identifier. E.g. NIS-code, NUTS-code, etc.

How to read in your data depends on the format, three recommended R-packages should cover mosts posibilities: 

* [readr](https://readr.tidyverse.org/): read/write plain text formats such as CSV, TXT, etc.
* [readxl](https://readxl.tidyverse.org/): read in Excel-files (.xlsx, .xls).
* [haven](https://haven.tidyverse.org/): read/write datasets from SAS, SPSS, Stata.

```{r}
# read in employment/unemployment numbers

# CSV
data_muni <- read_csv('data/muni_typology.csv', col_types = cols(.default = col_character()))

# Excel


# SAS


# Stata


```



# Join spatial data and data-of interest 

Recommended options:

1. general dataframe-join functions from dplyr: `left_join()`.
2. map-specific helper function from tmaptools: ``.

```{r, eval=FALSE}
# option 1 (dplyr): 
data = left_join(map_data, data_of_interest, by = "identifier")
data = left_join(map_data, data_of_interest, by = c("map_identifier" = "data_identifier"))
```

```{r, eval = FALSE}
# option 2 (maptools):

```

# Examples

## EX. Municipal socio-economic typology

```{r}
data("BE_ADMIN_MUNTY")
map_muni <- st_as_sf(BE_ADMIN_MUNTY)
data_muni <- read_csv('data/muni_typology.csv', col_types = cols(.default = col_character()))
muni <- left_join(map_muni, data_muni, by = c('CD_MUNTY_REFNIS' = 'gemeente_nis_code'))
```

```{r}
qtm(muni, fill = 'hoofdcluster_lbl')
```

## EX. Percentage of part-time workers in EU-countries

```{r, message=FALSE, warning=FALSE}
library(eurostat)
worktime_data <- get_eurostat('lfsi_pt_a') %>%
  filter(age == 'Y20-64', 
         worktime == 'TEMP', 
         sex == 'T', 
         time == '2017-01-01', 
         unit == 'PC_EMP')

# alternative, load saved eurostat-data:
#XXXX
#XXXX
#XXXX
#XXXX
```

```{r, message=FALSE, warning=FALSE}
map_data <- get_eurostat_geospatial(
  resolution = "60", # detail 
  nuts_level = "0") # NUTS 0-3

map_data <- st_crop(map_data, c(xmin=-10, xmax=45, ymin=36, ymax=71))
```

```{r, message=FALSE, warning=FALSE}
worktime <- left_join(map_data, worktime_data, by = c('CNTR_CODE' = 'geo'))
```

```{r}
qtm(worktime, fill = 'values', fill.title = 'Percentage part-time')
```


```{r}
# countrycode
```

