---
title: "Adding data"
output: 
    html_document:
      df_print: tibble
      toc: yes
      toc_float: yes
---

We have a (spatial) dataframe with the blank map-information, now we want to add our data-of-interest. Two steps:

1. Read in data-of-interst into R.
2. Join data-of-interest with the spatial dataframe.


```{r}
library(BelgiumMaps.StatBel)
library(tmap)
library(tmaptools)
library(sf)
library(dplyr)
library(readr)
library(readxl)
library(haven)
```

# Read in data-of-interest

Crucial: have a variable/column in your that contains the appropriate spatial identifier. E.g. NIS-code, NUTS-code, etc.

How to read in your data depends on the format, three recommended R-packages should cover mosts posibilities: 

* [readr](https://readr.tidyverse.org/): read/write plain text formats such as CSV, TXT, etc.
* [readxl](https://readxl.tidyverse.org/): read in Excel-files (.xlsx, .xls).
* [haven](https://haven.tidyverse.org/): read/write datasets from SAS, SPSS, Stata.

```{r}
# read in Belfius socio-economic municipality typology

data_muni <- read_excel('data/muni_typology.xlsx')    # Excel
data_muni <- read_sas('data/muni_typology.sas7bdat')  # SAS
data_muni <- read_dta('data/muni_typology.dta')       # Stata
data_muni <- read_sav('data/muni_typology.sav')       # SPSS

# CSV
data_muni <- read_csv( 
  file = 'data/muni_typology.csv', 
  col_types = cols(.default = col_character()))       # explicit: all strings
```


# Join spatial data and data-of interest 

Recommended options:

1. general dataframe-[join functions from dplyr](https://stat545.com/bit001_dplyr-cheatsheet.html): `left_join()`.
2. map-specific [helper function from tmaptools](https://rdrr.io/cran/tmaptools/man/append_data.html): `append_data()`.

```{r, eval=FALSE}
# option 1 (dplyr): 
library(dplyr)
data <- left_join(map_data, data_of_interest, by = "identifier")
data <- left_join(map_data, data_of_interest, by = c("map_identifier" = "data_identifier"))
```

```{r, eval = FALSE}
# option 2 (maptools):
library(tmaptools)
data <- append_data(map_data, data_of_interest, 
                    key.shp = "map_identifier", key.data = "data_identifier")
```

# Examples

## Ex. Municipal socio-economic typology

```{r}
# load map data
data("BE_ADMIN_MUNTY")
map_muni <- st_as_sf(BE_ADMIN_MUNTY)

# load data-of-interest
data_muni <- read_csv('data/muni_typology.csv', col_types = cols(.default = col_character()))

# join with left_join()
muni <- left_join(map_muni, data_muni, by = c('CD_MUNTY_REFNIS' = 'gemeente_nis_code'))
```

```{r}
qtm(muni, fill = 'hoofdcluster_lbl', fill.title = 'Socio-economic cluster')
```

## Ex. Part-time workers in the EU

```{r}
# Read Eurostat data on percentage of part-time employment 
worktime_data <- read_excel('data/eurostat_workingtime_2017.xlsx')
```


```{r, message=FALSE, warning=FALSE}
# alternatively, fetch this data directly:
library(eurostat)
worktime_data <- get_eurostat('lfsi_pt_a') %>%
  filter(age == 'Y20-64', 
         worktime == 'TEMP', 
         sex == 'T', 
         time == '2017-01-01', 
         unit == 'PC_EMP')
```

```{r, message=FALSE, warning=FALSE}
# load EU NUTS0 (country) map data directly from Eurostat
map_data <- get_eurostat_geospatial(
  resolution = "60", # detail 
  nuts_level = "0") # NUTS 0-3

# crop map data to "mainland" EU
map_data <- st_crop(map_data, c(xmin=-10, xmax=45, ymin=36, ymax=71))
```

```{r, message=FALSE, warning=FALSE}
# join map and workingtime data in one dataframe
worktime <- left_join(map_data, worktime_data, by = c('CNTR_CODE' = 'geo'))
```

```{r}
qtm(worktime, fill = 'values', fill.title = 'Percentage part-time')
```

**Tip**: Use the R [countrycode package](https://github.com/vincentarelbundock/countrycode#r-countrycode) to convert names, codes, etc. before merging. Contains 30+ different country coding schemes, and to 600+ variants of country names in different languages and formats.

```{r}
library(countrycode)
countrycode(worktime$NUTS_ID, 'eurostat', 'ecb') # official ECB code
countrycode(worktime$NUTS_ID, 'eurostat', 'un.name.fr') # full UN name in FR
```

