---
title: "Tweak"
output: 
    html_document:
        toc: TRUE
        toc_float: TRUE
---

**Two** types of data-manipulation: 

1. spatial data-manipulations 
2. 'regular' data-manipulations

... but same syntax and in **one** dataframe.

PS: always an option, do (2) in SAS, Stata, etc. before reading and joining data (previous step.

```{r, message=FALSE, warning=FALSE}
library(BelgiumMaps.StatBel)
library(sf)
library(tmap)
library(dplyr)
```


# "Regular" data-manipulations

proportion of unemployed instead of 

# Spatial data-manipulations

Already mentioned spatial **cropping**. Also potentially useful for thematic maps: spatial **union**, spatial **aggregation**. Further (complex) operations, depending on need.

## Ex.: Crop provinces around Brussels Airport


FYI, fancy example: crop map Dublin airport.


## Ex.: Join Brussels to Flemish provinces

```{r}
data("BE_ADMIN_PROVINCE")
prov <- st_as_sf(BE_ADMIN_PROVINCE)
qtm(prov)
```

```{r}
data("BE_ADMIN_REGION")
region <- st_as_sf(BE_ADMIN_REGION)
bxl <- region %>% filter(TX_RGN_DESCR_NL == "Brussels Hoofdstedelijk Gewest")
qtm(bxl)
```




## Ex. Aggregate muncipalities to arrondissements


```{r}
# aggr <- muni %>%
#   #group_by(TX_RGN_DESCR_NL, hoofdcluster_lbl) %>%
#   group_by(hoofdcluster_lbl) %>%
#   # filter(hoofdcluster_lbl == 'Sterk verstedelijkt') %>%
#   tally()
```


```{r}
# qtm(aggr, fill = 'hoofdcluster_lbl')
```


# EX. aggregate municipal income 



# EX. heritage plans in Leuven


```{r, message=FALSE}
library(sf)
library(tmap)
library(tmaptools)  # provides bb() for bounding box
library(rgdal)      # provides readOGR()
library(mapview)
```

```{r}
heritage <- st_as_sf(readOGR('data/heritage_plans/', layer = 'heritage_plans'))
st_crs(heritage) <- 31370
```

```{r}
# Get a bounding box around Leuven
leuven_bb <- bb('Leuven', projection = '31370') # set projection to BE Lambert 72 https://epsg.io/31370
leuven_bb
```

```{r}
#quick thematic map with a bounding box around Leuven
#qtm(heritage) # compare full map
#qtm(heritage, bbox = 'Leuven')

qtm(heritage, bbox = leuven_bb)
```

```{r}
heritage_leuven <- crop_shape(heritage, leuven_bb)
heritage_leuven
```


```{r}
mapview(heritage_leuven)
```


```{r}
qtm(heritage_leuven)
```

```{r}
library(dplyr)
library(stringr)

heritage_leuven <- heritage_leuven %>%
  mutate(start_year = str_sub(STARTDATUM, start = 1, end = 4))
```


```{r}
qtm(heritage_leuven, fill = 'start_year')
```
