---
title: "Plotting maps"
output: 
    html_document:
        toc: TRUE
        toc_float: TRUE
---

Recommended packages:

* [tmap](https://github.com/mtennekes/tmap#tmap-thematic-maps-in-r): special-purpose package for *t*hematic *map*s in R. [**focus**]
* [ggplot](https://ggplot2.tidyverse.org/): most popular package for datavis in R, recent and growing support for maps.
* [leaflet](https://rstudio.github.io/leaflet/): flexible, general purpose package for interactive maps.
* [mapview](https://r-spatial.github.io/mapview/): quick, single-function to interactively view spatial data(frames).

Learn more:

* [Stackoverflow](https://stackoverflow.com).
* Documentation tmap (F1).
* Books: "[An Introduction to R for Spatial Analysis and Mapping](https://books.google.be/books?id=iwJ6DwAAQBAJ) (2018)", "[Geocomputation with R](https://geocompr.robinlovelace.net/]) (full-text)", "[Data Visualization. A practical introduction](https://socviz.co/)" (full text).
* [Datacamp courses](https://www.datacamp.com/).


```{r, message=FALSE, warning=FALSE}
library(BelgiumMaps.StatBel)
library(mapview)
library(sf)
library(tmap)
library(readr)
library(dplyr)
library(ggplot2)
```


# Thematic maps with tmap

In an nutshell:

* **Q**uick **t**hematic **m**ap: `qtm()`
* Build map in layers with lots of control and options: `tm_shape()`, `tm_fill()`, `tm_borders()`, etc.
* Cherry-on-top: switch between interactive and static plotting with `tmap_mode('view')` and `tmap_mode('plot')`.

Some resources:

* [Getting started with tmap](https://cran.r-project.org/web/packages/tmap/vignettes/tmap-getstarted.html)
* https://geocompr.robinlovelace.net/adv-map.html
* Background and walkthrough: [Tennekes, M., 2018, tmap: Thematic Maps in R, Journal of Statistical Software, 84(6), 1-39](https://www.jstatsoft.org/article/view/v084i06).

Already showed some examples, here extra focus on:

* "Layered" building of maps.
* Binning and color scales
* Multiple (styled) borders
* Multiple maps

Tip: for _final_ tweaks, save map as SVG in [Inkscape](https://inkscape.org/).


```{r}
# load boundary for entire BE
data("BE_ADMIN_BELGIUM")
be <- st_as_sf(BE_ADMIN_BELGIUM)

# load municipal boundaries
data("BE_ADMIN_MUNTY")
munip_map <- st_as_sf(BE_ADMIN_MUNTY)

# load fiscal income data on municipal level
munip_data <- read_csv(
  file = 'data/fiscal_incomes_2016.csv', 
  col_types = cols(
    munip_label = col_character(),
    munip_nis = col_character(),
    n_inhabitants = col_integer(),
    income_mean = col_integer() ))

# add map and income data together on muncipal level
munip <- left_join(
  munip_map, munip_data, 
  by = c('CD_MUNTY_REFNIS' = 'munip_nis'))
```


## Building maps in layers

```{r}
tm_shape(munip) +
  tm_borders()
```

```{r}
tm_shape(munip) +
  tm_fill(col = 'income_mean', title = 'Mean income (2016)')
```

```{r}
tm_shape(munip) +
  tm_borders() +
  tm_fill(col = 'income_mean', title = 'Mean income (2016)')
```

```{r}
tm_shape(munip) +
  tm_borders(col = 'white', lwd = 0.3) +
  tm_fill(col = 'income_mean', title = 'Mean income (2016)')
```

```{r}
tm_shape(munip) +
  tm_borders(col = 'white', lwd = 0.3) +
  tm_fill(col = 'income_mean', title = 'Mean income (2016)', legend.hist = TRUE) +
  tm_legend(legend.outside = TRUE)
```

```{r}
 tm_shape(munip) +
  tm_borders(col = 'white', lwd = 0.3) +
  tm_fill(col = 'income_mean', title = 'Mean income (2016)', legend.hist = TRUE) +
  tm_legend(legend.outside = TRUE) +
  tm_layout(frame = FALSE) +
  tm_credits('Source: Statbel, Fiscale statistiek van de inkomsten 2005-2016', position = 'left')
```

```{r, message=FALSE, warning=FALSE}
 tm_shape(munip) +
  tm_borders(col = 'white', lwd = 0.3) +
  tm_fill(col = 'income_mean', title = 'Mean income (2016)', legend.hist = FALSE) +
  tm_legend(legend.outside = FALSE) +
  tm_layout(frame = FALSE) +
  tm_credits('Source: Statbel, Fiscale statistiek van de\ninkomsten 2005-2016', position = 'center') +
  tm_style_grey(legend.format = list(text.separator= "-")) + 
  tm_logo("data/hiva_logo_400x400.png")
```

## Add layers using objects

```{r}
be_income <- tm_shape(munip) +
  tm_borders(col = 'white', lwd = 0.3) +
  tm_fill(col = 'income_mean', title = 'Mean income (2016)', legend.hist = FALSE) +
  tm_legend(legend.outside = FALSE)

be_income
```

```{r}
be_income_style <- be_income +
  tm_layout(frame = FALSE) +
  tm_credits('Source: Statbel, Fiscale statistiek van de\ninkomsten 2005-2016', position = 'center') +
  tm_style_grey(legend.format = list(text.separator= "-")) + 
  tm_logo("data/hiva_logo_400x400.png", height = 2) +     # scale down logo size a bit
  tm_layout(scale = 0.8)                                  # reduce over text-size

be_income_style
```


```{r}
save_tmap(be_income_style, 'output/be_income_muni_2016.png', width = 1920, height = 1080)
```


# Styled (multiple) borders

```{r}
data("BE_ADMIN_PROVINCE")
prov <- st_as_sf(BE_ADMIN_PROVINCE)
```


```{r}
be_income_prov <- be_income +
  tm_shape(prov) +
  tm_borders(col = 'black', lwd = .2)

be_income_prov
```

## Binning and color palettes

* Choice of binning intervals matters.
* Choise of color palette matters, cf. [colorbrewer](http://colorbrewer2.org): sequential, diverging, and qualitative color palettes.

```{r}
min(munip$income_mean)
max(munip$income_mean)
```


```{r}
tm_shape(munip) +
  tm_fill(col = 'income_mean', 
          title = 'Mean income (2016)',
          breaks = c(0, 20000, 25000, 30000)) +
  tm_borders(col = 'white', lwd = .5)
```



```{r}
tm_shape(munip) +
  tm_fill(col = 'income_mean', 
          title = 'Mean income (2016)',
          n=20, # detailed breaks!
          legend.hist = TRUE) +
  tm_borders(col = 'white', lwd = .5) +
  tm_legend(legend.outside = TRUE, legend.outside.position="right")
```

```{r}
be_income_div <- tm_shape(munip) +
  tm_fill(col = 'income_mean', 
          title = 'Mean income (2016)',
          n = 20,  
          palette = "RdBu", auto.palette.mapping = FALSE, # diverging pallette!
    legend.hist = TRUE) +
  tm_borders(col = 'white', lwd = .5) +
  tm_legend(legend.outside = TRUE, legend.outside.position="right") +
  tm_shape(be) +
  tm_borders(col = 'black', lwd = 0.3) +
  tm_layout(frame = FALSE)

be_income_div
```



# Multiple maps / facets

Facets: more info in [getting started document](https://cran.r-project.org/web/packages/tmap/vignettes/tmap-getstarted.html#facets) online.

```{r}
tm_shape(munip) +
  tm_borders(col = 'white', lwd = .5) +
  tm_fill(col = c("income_mean", 'n_inhabitants')) +
  tm_layout(legend.position = c('left', 'bottom'))
```


```{r}
m_be_income <- tm_shape(munip) +
  tm_borders(col = 'white', lwd = .5) +
  tm_fill(col = "income_mean", title = 'Mean income (2016)')

m_be_pop <- tm_shape(munip) +
  tm_borders(col = 'white', lwd = .5) +
  tm_fill(col = "n_inhabitants", title = "Population (2010)")

tmap_arrange(m_be_income, m_be_pop)
```


## tmap interactive mode

```{r, message=FALSE, warning=FALSE}
tmap_mode('view')
be_income_div
tmap_mode('plot')
```



# ggplot with geom_sf()

See chapter 7 of "[Data Visualisation. A Practical Introduction](http://socviz.co/maps.html#maps)".

```{r, message=FALSE, warning=FALSE}
library(eurostat)
worktime_data <- get_eurostat('lfsi_pt_a') %>%
  filter(age == 'Y20-64', 
         worktime == 'TEMP', 
         sex %in% c('M', 'F'), 
         time == '2017-01-01', 
         unit == 'PC_EMP')

map_data <- get_eurostat_geospatial(
  resolution = "60", # detail 
  nuts_level = "0") # NUTS 0-3
map_data <- st_crop(map_data, c(xmin=-10, xmax=45, ymin=36, ymax=71))

worktime <- left_join(map_data, worktime_data, by = c('CNTR_CODE' = 'geo')) %>%
  filter(!is.na(sex))
```

```{r, message=FALSE, warning=FALSE}
ggplot(worktime) +
  geom_point(aes(x = values, y = id, color = sex)) +
  facet_grid(~sex)
```


```{r}
p_eu_worktime <- ggplot(worktime) +
  geom_sf(aes(fill = values)) +
  facet_grid(~sex)

p_eu_worktime
```

```{r}
ggsave('output/eu_worktime_gender.png', p_eu_worktime, width = 12, height = 8)
```



# Interactive maps with leaflet 

Good [intro online](https://rstudio.github.io/leaflet/choropleths.html).

```{r, warning=FALSE, message=FALSE}
library(leaflet)

bins <- c(0, 5, 10, 15, 20, 25, 30)
pal <- colorBin("YlOrRd", domain = worktime$values, bins = bins)

eu_worktime_leaflet <- leaflet(worktime %>% filter(sex == 'F')) %>%
  addTiles() %>%
  addPolygons(fillColor = ~pal(values), color = "black", weight = 1, opacity = 1)

eu_worktime_leaflet
```

**Tip**: use `saveWidget()` from library [htmlwidgets](https://www.rdocumentation.org/packages/htmlwidgets/versions/1.3/topics/saveWidget) to save a (standalone) interactive map.

```{r, eval=FALSE}
library(htmlwidgets)
saveWidget(eu_worktime_leaflet, 'output/eu_worktime_leaflet.html', selfcontained = TRUE)
```

Leaflet is quite flexible, some further examples:

* [Interactive map in press-release](https://hiva.kuleuven.be/nl/nieuws/nieuwsitems/bruto-jobtoename-en-afname-in-beeld-voor-de-Belgische-regios-en-provincies) on unemployment-numbers HVIA-website.
* Quick [example showing sampling in statistical sectors](https://mhermans.net/files/tmp/map_sectoren_basic.html) for project-meeting.
* [Icons on a thematic map](https://rawgit.com/mhermans/rgeonotebooks/master/vdab_api_popup_map/vdab_api_popup_map.nb.html) (example with VDAB-offices).
* Plotting on [interactive historical maps](https://mhermans.net/post/mapping-leuvense-gangen/).
