---
title: "Tweak"
output: 
    html_document:
        toc: TRUE
        toc_float: TRUE
---

Recommended packages:

* [mapview](https://r-spatial.github.io/mapview/): quick, single-function to interactively view spatial data(frames).
* [leaflet](https://rstudio.github.io/leaflet/): flexible, general purpose package for interactive maps.
* [ggplot](https://ggplot2.tidyverse.org/): most popular package for datavis in R, recent and growing support for maps.
* [tmap](https://github.com/mtennekes/tmap#tmap-thematic-maps-in-r): special-purpose package for *t*hematic *map*s in R. [focus]

Learn more:

* Stackoverflow
* Documentation tmap
* Datacamp courses
* Online books, https://geocompr.robinlovelace.net/adv-map.html
* Offline books



```{r, message=FALSE, warning=FALSE}
library(BelgiumMaps.StatBel)
library(mapview)
library(sf)
library(tmap)
library(readr)
library(dplyr)
library(ggplot2)
```




# Themaric maps with tmap

In an nutshell:

* **Q**uick **t**hematic **m**ap: `qtm()`
* Build map in layers with lots of control and options: `tm_shape()`, `tm_fill()`, `tm_borders()`, etc.
* Cherry-on-top: switch between interactive and static plotting with `tmap_mode('view')` and `tmap_mode('plot')`.


* [Getting started with tmap](https://cran.r-project.org/web/packages/tmap/vignettes/tmap-getstarted.html)
* https://geocompr.robinlovelace.net/adv-map.html
* Background and walkthrough: [Tennekes, M., 2018, tmap: Thematic Maps in R, Journal of Statistical Software, 84(6), 1-39](https://www.jstatsoft.org/article/view/v084i06).

Already showed some examples, here extra focus on:

* Color scales
* Multiple (styled) borders
* Multiple maps

Tip: for _final_ tweaks, save map as SVG in [Inkscape](https://inkscape.org/).


```{r}
# load boundary for entire BE
data("BE_ADMIN_BELGIUM")
be <- st_as_sf(BE_ADMIN_BELGIUM)

# load municipal boundaries
data("BE_ADMIN_MUNTY")
munip_map <- st_as_sf(BE_ADMIN_MUNTY)

# load fiscal income data on municipal level
munip_data <- read_csv(
  file = 'data/fiscal_incomes_2016.csv', 
  col_types = cols(
    munip_label = col_character(),
    munip_nis = col_character(),
    n_inhabitants = col_integer(),
    income_mean = col_integer() ))

# add map and income data together on muncipal level
munip <- left_join(
  munip_map, munip_data, 
  by = c('CD_MUNTY_REFNIS' = 'munip_nis'))
```


## Building maps in layers

```{r}
tm_shape(munip) +
  tm_borders()
```

```{r}
tm_shape(munip) +
  tm_fill(col = 'income_mean', title = 'Mean income (2016)')
```

```{r}
tm_shape(munip) +
  tm_borders() +
  tm_fill(col = 'income_mean', title = 'Mean income (2016)')
```

```{r}
tm_shape(munip) +
  tm_borders(col = 'white', lwd = 0.3) +
  tm_fill(col = 'income_mean', title = 'Mean income (2016)')
```

```{r}
tm_shape(munip) +
  tm_borders(col = 'white', lwd = 0.3) +
  tm_fill(col = 'income_mean', title = 'Mean income (2016)', legend.hist = TRUE) +
  tm_legend(legend.outside = TRUE)
```

```{r}
 tm_shape(munip) +
  tm_borders(col = 'white', lwd = 0.3) +
  tm_fill(col = 'income_mean', title = 'Mean income (2016)', legend.hist = TRUE) +
  tm_legend(legend.outside = TRUE) +
  tm_layout(frame = FALSE) +
  tm_credits('Source: Statbel, Fiscale statistiek van de inkomsten 2005-2016', position = 'left')
```

```{r, message=FALSE, warning=FALSE}
 tm_shape(munip) +
  tm_borders(col = 'white', lwd = 0.3) +
  tm_fill(col = 'income_mean', title = 'Mean income (2016)', legend.hist = FALSE) +
  tm_legend(legend.outside = FALSE) +
  tm_layout(frame = FALSE) +
  tm_credits('Source: Statbel, Fiscale statistiek van de\ninkomsten 2005-2016', position = 'center') +
  tm_style_grey(legend.format = list(text.separator= "-")) + 
  tm_logo("data/hiva_logo_400x400.png")
```

## Add layers using objects

```{r}
be_income <- tm_shape(munip) +
  tm_borders(col = 'white', lwd = 0.3) +
  tm_fill(col = 'income_mean', title = 'Mean income (2016)', legend.hist = FALSE) +
  tm_legend(legend.outside = FALSE)

be_income
```

```{r}
be_income_style <- be_income +
  tm_layout(frame = FALSE) +
  tm_credits('Source: Statbel, Fiscale statistiek van de\ninkomsten 2005-2016', position = 'center') +
  tm_style_grey(legend.format = list(text.separator= "-")) + 
  tm_logo("data/hiva_logo_400x400.png", height = 2) +     # scale down logo size a bit
  tm_layout(scale = 0.8)                                  # reduce over text-size

be_income_style
```


```{r}
save_tmap(be_income_style, 'output/be_income_muni_2016.png', width = 1920, height = 1080)
```


# Styled (multiple) borders

```{r}
data("BE_ADMIN_PROVINCE")
prov <- st_as_sf(BE_ADMIN_PROVINCE)
```


```{r}
be_income_prov <- be_income +
  tm_shape(prov) +
  tm_borders(col = 'black', lwd = .2)

be_income_prov
```

## Binning and color palettes


Continuous, sequential, diverging.

```{r}
min(munip$income_mean)
max(munip$income_mean)
```


```{r}
tm_shape(munip) +
  tm_fill(col = 'income_mean', 
          title = 'Mean income (2016)',
          breaks = c(0, 20000, 25000, 30000)) +
  tm_borders(col = 'white', lwd = .5)
```



```{r}
tm_shape(munip) +
  tm_fill(col = 'income_mean', 
          title = 'Mean income (2016)',
          n=20, # detailed breaks!
          legend.hist = TRUE) +
  tm_borders(col = 'white', lwd = .5) +
  tm_legend(legend.outside = TRUE, legend.outside.position="right")
```

```{r}
be_income_div <- tm_shape(munip) +
  tm_fill(col = 'income_mean', 
          title = 'Mean income (2016)',
          n = 20,  
          palette = "RdBu", auto.palette.mapping = FALSE, # diverging pallette!
    legend.hist = TRUE) +
  tm_borders(col = 'white', lwd = .5) +
  tm_legend(legend.outside = TRUE, legend.outside.position="right") +
  tm_shape(be) +
  tm_borders(col = 'black', lwd = 0.3) +
  tm_layout(frame = FALSE)

be_income_div
```



# Multiple maps / facets

https://cran.r-project.org/web/packages/tmap/vignettes/tmap-getstarted.html#facets


```{r}
tm_shape(munip) +
  tm_borders(col = 'white', lwd = .5) +
  tm_fill(col = c("income_mean", 'n_inhabitants')) +
  tm_layout(legend.position = c('left', 'bottom'))
```


```{r}
m_be_income <- tm_shape(munip) +
  tm_borders(col = 'white', lwd = .5) +
  tm_fill(col = "income_mean", title = 'Mean income (2016)')

m_be_pop <- tm_shape(munip) +
  tm_borders(col = 'white', lwd = .5) +
  tm_fill(col = "n_inhabitants", title = "Population (2010)")

tmap_arrange(m_be_income, m_be_pop)
```


## tmap interactive mode

```{r, message=FALSE, warning=FALSE}
tmap_mode('view')
be_income_div
tmap_mode('plot')
```



# ggplot with geom_sf()

```{r, message=FALSE, warning=FALSE}
library(eurostat)
worktime_data <- get_eurostat('lfsi_pt_a') %>%
  filter(age == 'Y20-64', 
         worktime == 'TEMP', 
         sex %in% c('M', 'F'), 
         time == '2017-01-01', 
         unit == 'PC_EMP')

map_data <- get_eurostat_geospatial(
  resolution = "60", # detail 
  nuts_level = "0") # NUTS 0-3
map_data <- st_crop(map_data, c(xmin=-10, xmax=45, ymin=36, ymax=71))

worktime <- left_join(map_data, worktime_data, by = c('CNTR_CODE' = 'geo')) %>%
  filter(!is.na(sex))
```

```{r, message=FALSE, warning=FALSE}
ggplot(worktime) +
  geom_point(aes(x = values, y = id, color = sex)) +
  facet_grid(~sex)
```


```{r}
p_eu_worktime <- ggplot(worktime) +
  geom_sf(aes(fill = values)) +
  facet_grid(~sex)

p_eu_worktime
```

```{r}
ggsave('output/eu_worktime_gender.png', p_eu_worktime, width = 12, height = 8)
```



# Interactive maps with leaflet 

https://rstudio.github.io/leaflet/choropleths.html

saving leaflet 

```{r, warning=FALSE, message=FALSE}
library(leaflet)

bins <- c(0, 5, 10, 15, 20, 25, 30)
pal <- colorBin("YlOrRd", domain = worktime$values, bins = bins)

leaflet(worktime %>% filter(sex == 'F')) %>%
  addTiles() %>%
  addPolygons(fillColor = ~pal(values), color = "black", weight = 1, opacity = 1)
```

* Unemployment dynam-project HVIA-site
* statistical sectors
* Example VDAB-offices
* Historical maps
* 
