---
title: "Show & tell: thematic maps with R"
---

# Schedule

* Monday 11/02
    - Intro-presentation.
    - Examples [step 1, 2] getting maps & adding data.
* Thursday 14/02
    - Short recapitulation intro-presentation.
    - Examples [step 3, 4] manipulating data & plotting.
    - [depending on timing] Q&A, own examples / data.

# Intro-presentation

1. (Dis)advantages of R
2. Why spatial datavis. in R is become increasingly interesting
    - tidyverse
    - spatial features
    - one workflow for spatial & regular data
3. Recommended packages for thematic maps
    - tmap
    - ggplot / leaflet / mapview
    - supporting packages
4. Outline of worked examples / approach
5. Expectations? Questions?

# Setup

* [Download](https://cran.r-project.org/bin/windows/base/) and install R.
* [Download](https://www.rstudio.com/products/rstudio/download/#download) and install Rstudio.
* Run commands below to install required R-packages.

```{r, eval=FALSE}
install.packages(c('eurostat', 'rworldmap', 'rgdal', 'sf', 'osmdata', 'tmap', 'tmaptools', 'countrycode', 'ggplot2'))
```

# R-syntax in a nutshell (enough to follow along)

```{r, eval=FALSE}
# Objects, functions, combinations of functions

obj_a <- something            # create/assign something to "object A"
obj_a                         # show object A

obj_b <- do_something(obj_a)  # change object A, creating object B (function)
 
# %>% "pipe" object A through multiple functions, resulting in object B
obj_b <- obj_a %>%
  do_something() %>%
  do_something_else() %>%
  do_another_thing()

# + add things to object A, resulting in object B
obj_b <- obj_a +
  add_something() +
  add_something()
```
